### 2023-04-05 reachable-resilience.tex
up

\section{State resilience}

Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$. But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap post^*(s_0)$, there must exist a path to $\Safe$. 
%
%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.

The three previous problems become:

\problemx{State-resilience problem (SRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
% 		an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
% 		{$ ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe, ~ s' \rightarrow^{*} s''$ ?\newline}
{$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}

\problemx{$k$-state-resilience problem (kSRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
%

\problemx{bounded-state-resilience problem (BSRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient. ?\newline}

%\textcolor{blue}{
%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
%}

Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.

As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downard closed and upward closed. We will consider here the cases where 
$\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
and where 
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.

% \mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}

\subsection{Case: $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$}

Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
This stems from the fact that $t$-liveness is undecidable for WSTS and 
is reducible to the state resilience problem.

\begin{proposition}\label{reductions}
In WSTS, $t$-liveness is reducible to the state resilience problem.
%	home state and reachability
\end{proposition}

\begin{proof}
We reformulate $t$-liveness as the following formula
% $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
$$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$  
where
$U_t=\uparrow \pred(t)$
is the upward closure of the preconditions to use transition $t$.  
The problem reduces itself to the state resilience problem
where $\Safe = U_t$ and $\Bad = S \setminus \Safe$.
\end{proof}

To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. The following theorem is a adjustment ofTheorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.

\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
The  Reset VASS zero reachability problem can be reduced to the $t$-liveness problem.
\end{theorem}

\iffalse
% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
\begin{proof}
See Appendix~\ref{appendix} for the complete proof.
\end{proof}
\fi

Since the Reset VASS zero reachability problem is undecidable, the reduction implies 
% the following:
%
\iffalse
\begin{corollary}
Reset VASS liveness is undecidable.
\end{corollary}

Since moreover liveness can be reduced to 
$t$-liveness, we deduce:
\fi
%
%\begin{corollary}
$t$-liveness is undecidable for Reset VASS.
%\end{corollary}

\iffalse

\begin{theorem}
The downward reachability problem is reducible to the state resilience problem.
\end{theorem}

\begin{proof}
Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
Take $\Safe = \emptyset$ and $\Bad = D$.
The state resilience problem on $\Safe,\Bad$ asks
whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
If the answer is negative then $s \to^* \Bad$,
else $s \not\to^* \Bad$.
\end{proof}

\fi

Since $t$-liveness is undecidable for Reset VASS that are WSTS with strong compatibility, from Proposition~\ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

\begin{theorem}\label{srp up down}
{\sc State-resilience},
{\sc Bounded-state-resilience} and
{\sc $k$-state-resilience}
are undecidable for strongly compatible WSTS with effective pred-basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

\begin{proof}
{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since $t$-liveness is undecidable.

% \begin{proposition}
% \end{proof}

% \begin{proposition}
In WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
reducible to {\sc $k$-state-resilience}:
% \end{proposition}
% 
% \begin{proof}
since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
and there exists $n_0 \in \N$ such that 
$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
We compute 
$n_0$, then iteratively check whether $k$-state resilience hold for $k$ from $0$ to $n_0$.  
% \end{proof}
Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
% \end{proposition}
%
% \begin{proof}
since 
$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
Hence the %equivalence, and
 undecidability of {\sc Bounded-state-resilience}
and 
 {\sc $k$-state-resilience}.
%
%
\iffalse
% \begin{corollary}\label{bsrp up down}
Bounded state-resilience and 
$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
% \end{corollary}
\fi
%
%\begin{proof} 
%  From the two reductions above. 
\end{proof}

On the positive side, let us recall a result about state-resilience decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).

\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

We may immediately generalyse this last result by strengthening to \emph{unbounded} state-resilience. The proof is essentially the same than the previous one.

\begin{corollary}\label{postcomputable}
{\sc state-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{corollary}

\begin{proof}
%From Fact~\ref{stop condition},
Since $\mathscr{S}$ is a WSTS there exists $n_0 \in \N$ such that
$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute this $n_0$ by iteratively computing 
%\alain{mauvaises notations: $k, k_m, Safe^k,...$ n'a aucun sens...}
$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
returning $n$ if that is the case.
Then, because $n_0$-state-resilience is decidable, 
checking $\uparrow post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
and state-resilience is decidable.
\end{proof}

The proof of Theorem~\ref{ref ozkan} rely on the computability of $\uparrow \post^*(s)$ and on the following lemma.

\begin{lemma}\label{Lemma intersection}
Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
\end{lemma}

\begin{proof}
Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
In the other direction,
since $A \subseteq {\uparrow A}$, the inclusion
$({\uparrow  A}) \cap D \subseteq U$ implies
$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
\end{proof}

\iffalse
\alain{definir downward reachability problem.....
downward-closed problem given a state $s$ of a WSTS
% in the regarded class 
with strong compatibility 
and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
\fi

The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.

\problemx{Downward-reachability problem}
{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
\subseteq S$.}
{$\exists s' \in D ~ s \to^* s'$? \newline}

%\"Ozkan
\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
For finite-branching WSTS%with strong compatibility
, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
%i 			.e. given a state $s$ of a WSTS
% in the regarded class 
%with strong compatibility 
%and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
\end{proposition}

% \alain{rappeler l'idée de la preuve}
The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
$B_{\uparrow \post^*(s)} \cap D = \emptyset$, that is equivalent to $\post^*(s)\cap D = \emptyset$ by
Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed sets
$U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
Decidability of downward reachability leads to the decidability of the following stop condition:
asking whether $S \setminus U_n$ is reachable from $s$.

The characterization is used to show that VASS are $\uparrow \post^*$-effective. Since the downward-reachability problem is decidable for VASS, we deduce the following:

\begin{proposition}
Let $V= (Q,T)$ be a $d$-VASS and $v \in \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(v)$.
\end{proposition}
%\alain{mauvaises notations, la tradition est que P est l'ensemble des places...les états/markings en minuscules,....}

\iffalse
\begin{proof}
By Proposition~\ref{post*}, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
Let us now show that the downward-reachability problem is decidable for Petri nets.
Let us consider a downward closed set $D$ in a Petri net.
Let us consider $B$ a basis of the upward-closed set $S \setminus D$.

$D$ is  defined by ... states where your number of counter is bounded by a constant (deduce this from ... having to be smaller than elements of the basis $B$), and others where it is unbounded. You have $m$ counters where it is bounded by $n$
then you have $m \times n$ possibilities for the values in these counters
you have unlimited possibilities for the values of the other counters,
but that is where the submarking problem comes in handy,
i.e. you ask the submarking problem for all of the $m \times n$ possibilities.

So downward-set reachability is decidable for PN.
\end{proof}
\fi

It is well-known that 
% Petri nets are WSTS with strong  compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
% Since 
VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
Hence state-resilience is decidable for %both Petri nets and 
VASS.
However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:

\begin{proposition}
There exist classes of WSTSs with strong 
%(upward)
 compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
\end{proposition}

\begin{proof}
Let us show that Reset VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
\iffalse
Then, one would be able to decide whether an element $m \in Min(\uparrow \post^*)$ is reachable.
%in counter $i$ 
by examining if there is %some vector $v$ in the basis such that $v(i)=0$
$0$ in the basis%
. But reachability of $0$  %in a counter $i$ 
is undecidable for Reset VASS. 
% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\fi
Then, {\sc State-resilience} would be decidable for Reset VASS, which leads to a contradiction.
\end{proof}

%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.

%
%si on garde strong mais on enlève effective basis of $post^*$:
%
%\begin{proposition}(à prouver)
%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong compatibility.
%\end{proposition}
%
%\begin{proof}
%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
%\end{proof}

Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.

\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
{\sc state-resilience} is decidable for 
%(not necessarily strong)
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

\begin{proof}
Let $B_s$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
and $\Bad$ a downward-closed set given by its finite set of (maximal) ideals.
By applying Lemma~\ref{Lemma intersection} twice, we obtain

$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_s \cap \Bad \subseteq \pred^*(\Safe)$$

Since $B_s$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_s \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
\end{proof}

% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
%
% Well, maybe their version of resilience but resilience with * seems fine.
%

However when removing strong compatibility, precision is lost.
Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
 $\uparrow post^* (s) \cap S \not\subseteq \pred(\Safe)$,
despite having 
$post^* (s) \cap S \subseteq \pred( \Safe)$.
In such a case the algorithm from
\cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold,
which is incorrect.
% An example of this is ...

Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
{\sc bounded-state-resilience}.

\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}

Unfortunately the problem is undecidable in general.

\begin{theorem}\label{srp down up}
{\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
when
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$.
\end{theorem}

\begin{proof}
Consider Reset VASS
in the case $\Safe$ contains only the minimal element $Min(S) = 0$ %$0$ 
and $\Bad = S \setminus \Safe$, 
it follows the state resilience problem is equivalent 
to the problem of whether $0$ can be visited infinitely often. 
% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
Hence the undecidability.
\end{proof}

Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
%
For instance one can write the following lemma:

\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
Then $A \cap U \subseteq D$  iff $ (\downarrow  A) \cap U \subseteq D$.
\end{lemma}

\begin{proof}
Let us suppose that $A \cap U \subseteq D$. Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$. Let us show that $({\downarrow A}) \cap U \subseteq {\downarrow (A \cap U)}$. Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$. Since $x \in U$ and $U$ is downward-closed, we also have $a \in U$. Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$. In the other direction, since $A \subseteq {\downarrow A}$, the inclusion $({\downarrow  A}) \cap U \subseteq D$ implies $A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
\end{proof}

In the case of a WSTS with downward compatiblitity, not necessarily strong,
then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
Lemma~\ref{Lemma intersection 2} can be used to show that
if $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$,
then
$\post^*(s) \cap \Bad \subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$.

\begin{theorem}
{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{theorem}

\begin{proof}
In order to decide whether state-resilience hold, we execute two procedures in parallel,
one looking for a resilience certificate and one looking for a non-resilience certificate.
Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
Notice that, by standard monotonicity, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.

Procedure 1 stops when it finds an invariant $D$ such that
$D  \cap \Bad \subseteq \pred^*(\Safe)$. 
Indeed
$D  \cap \Bad \subseteq  \pred^*(\Safe)$ implies
$\downarrow \post^*(s) \cap \Bad \subseteq  \pred^*(\Safe)$
since $ \downarrow \post^*(s)  \subseteq D$.

The second procedure iteratively computes
$post^{\leq n}(s) \cap \Bad$
until it finds an element
not in $\pred^*(\Safe)$.
\end{proof}

% \mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of $\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}

Strong downward compatibility implies furthermore the decidability
of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.

\begin{proposition}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{proposition}

\subsection{Synthesis}

Results in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:

\begin{center}
\begin{tabular}{ | l | c | c | c | c |}
\hline  Hypothesis & strong upward compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
 \end{tabular}
\end{center}

\noindent
Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
\iffalse
(by symmetry) for downward compatible WSTS (but not necessarily upward compatible):

\begin{center}
\begin{tabular}{ | l | c | c | c | c |}
\hline  Hypothesis & strong downward compatibility ~ & $\downarrow post^*$ effective & both  \\ \hline
   SRP & Undecidable  & Decidable   & Decidable \\ \hline
   BSRP & Undecidable  &  ??  & Decidable \\ \hline
      kSRP & Undecidable  & ?? & Decidable  \\ \hline
 \end{tabular}
\end{center}

\mathieu{Montrer que leurs hypothèses Lossy + $\bot$-bounded + strong upward compatibility c'est plus fort que just strong downward compatibility et $\downarrow \post^*$ effective ?}

Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt
\fi

\begin{center}
\begin{tabular}{ | l | c | c | c | c |}
\hline  Hypothesis & $\emptyset $ & downward compatible & strongly downward compatible  \\ \hline
   SRP & Undecidable (Thm~\ref{srp down up})  & Decidable   & Decidable \\ \hline
   BSRP & ??  &  ??  & Decidable \\ \hline
      kSRP & ??  & ?? & Decidable  \\ \hline
 \end{tabular}
\end{center}