### 2023-06-08 application.tex
up

\section{Computing the maximal resilient subsystem}

Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.

\begin{definition}
We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
\end{definition}

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

\begin{theorem}{}
Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
\end{theorem}

\begin{proof}
Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
\end{proof}
%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
%
%
%\noindent
%{\bf Vector addition system with states}
%
Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}. 
% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e. $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
%
Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
%
\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

\noindent

Let  $V$ be a $d$-VASS, $X \subseteq \mathds{N}^d$ be an upward-closed set and $B_X$ its minimal basis: $X=\uparrow B_X$. The number of elements of $min(\pred^*(\uparrow B_X))$ and the size of the minimal elements in $min(\pred^*(X))$ have been studied in
\cite{DBLP:conf/rp/BozzelliG11}: both sizes are bounded by $2^{2^n}$ where $n$ is the size of $V$.

\begin{theorem}{}
Let $V$ be a VASS and $X$ be an upward-closed set. The maximal $X$-resilient subsystem of $V$ is a computable VASS of size $2^{2^n}$ where $n$ is the size of $V$.
\end{theorem}

We may use the previous Theorem with different upward-closed sets $X$. Let us first recall the definition of 4 types of configurations used in \cite{DBLP:journals/acta/ValkJ85}.

\begin{definition}
Let $V = (Q,T)$ be a $d$-VASS, $\hat{T} \subseteq T$, and $m$ a configuration in $Q \times \N^d$.
\begin{itemize}
\item  $m$ is \em{$\hat{T}$-blocked} if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
\item $m$ is \em{dead} if the reachability tree from $m$ is finite.
\item $m$ is \em{bounded} if the reachability set from $m$ is finite.
\item $m$ is \em{$\hat{T}$-continual} if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
\end{itemize}
\end{definition}

These 4 questions are decidable: $m$ is $\hat{T}$-blocked iff the preconditions of each transition in $\hat{T}$ is not coverable from $m$.
$m$ is dead iff $V$ with $m$ as initial configuration terminates. $m$ is bounded iff $V$ with $m$ as initial configuration is bounded.
Hence, the first two problems reduce to coverability which is EXPSPACE. The third problem is also well-known in EXPSPACE. The fourth problem can be decided by analysing circuits in the coverability graph \cite{DBLP:journals/acta/ValkJ85} but it can be also shown in EXPSPACE following techniques exposed in \cite{DBLP:journals/jcss/Demri13}.

Let us consider the following four associated upward-closed sets.

\begin{itemize}
\item  NotBlocked($\hat{T})= \{ m \in Q \times \N^d \mid m$ is not $\hat{T}$-blocked\}.
\item NotDead = $\{ m \in Q \times \N^d \mid m$ is not dead\}.
\item UnBounded = $\{ m \in Q \times \N^d \mid m$ is not bounded\}.
\item Continual($\hat{T}$) = $\{ m \in Q \times \N^d \mid m$ is $\hat{T}$-continual\}.
\end{itemize}

Let $(V,v_0)$ be a VASS with an initial state and $X$ be any of the 4 previous sets. Valk and Jantzen showed that the maximal subsystem of $(V,v_0)$ such that its reachability set $\post^*(V,v_0)$ is included in $X$ is still a VASS $(V_X,v_0)$ that can be constructed with the coverability graph \cite{DBLP:journals/acta/ValkJ85} whose complexity is Ackermanian. Let us remark that the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$, the maximal $X$-resilient VASS of $V$ is also a computable VASS $V_X$ and that $V_X$ is independant of the initial state $v_0$.

\section{Resilience for VASS}

\begin{proposition}
{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
\end{proposition}

\begin{theorem}{}
Resilience is decidable for VASS when $\Safe$ and $\Bad$ are semilinear sets.
\end{theorem}

\begin{proof}
When asking about {\sc Resilience}, in context, there are no true ``neutral'' configurations.
Indeed, from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
$V$, $\Safe$ and $\Bad$ if and only if {\sc Resilience} holds for
$V'$, $\Safe'$ and $\Bad'$. For the construction, $V'$ will have two copies of the control states of $V$; $Q_{A}$ and $Q_{B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp. $\Bad$. 
From a configuration with state in $Q_{A}$ you go to $Q_B$ when you are in a $\Bad$ place
and from a configuration with state in $Q_{B}$ you go to $Q_{A}$ when you are in a $\Safe$ place.
$\Safe'$ is $Q_A \times \N^d$ and $\Bad'$ is $Q_B \times N^d$.

Without loss of generality hence let us consider only the case when $\Safe$ is semilinear
and $\Bad = S \setminus \Safe$ is semilinear too.
{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).\]
Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
$\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
{\sc Resilience} is decidable.
\end{proof}

Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logics. Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.

\begin{theorem}{}
Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
\end{theorem}

\begin{proof}
We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear. Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
\end{proof}

Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.

\begin{theorem}{}
Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
\end{theorem}

\begin{proof}
The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
% For Z-VAS we observe that we are asking for inclusion between linear sets. 
% Again let us without loss of generality hence let us consider only the case when $\Safe$ is semilinear and $\Bad = S \setminus \Safe$ is semilinear too.
\mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
\end{proof}

Continuous VASS \cite{DBLP:journals/tocl/BlondinFHH17} are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.

\begin{theorem}{}
Resilience is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
the rationals with addition and order.
\end{theorem}

\begin{proof}
The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also in EXPSPACE. 
%			\alain{and may be in coNP-complete. Recall that reachability is P-complete for continuous VASS}
\end{proof}

\begin{figure}

\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad 
		& $\uparrow$~ $\downarrow$~ 
		 & $\downarrow$~ $\uparrow$~

 \\ \hline
   RP  
   	& Decidable (Thm~\ref{down-up})  
   		 & ??

    \\ \hline
   BRP  
   &  Decidable (Thm~\ref{down-up}) 
   		 & Decidable  ?

    \\ \hline
      kRP  
      & Decidable (Thm~\ref{down-up}) 
      		& Decidable ?

       \\ \hline
   sRP  
   	& Decidable (Thm~\ref{post srp})
   		 & ??

    \\ \hline
   BsRP  
   &  Decidable (Thm~\ref{ref ozkan})
   		 & ??

    \\ \hline
      ksRP   
      & Decidable (Thm~\ref{ref ozkan})
      		& ??

       \\ \hline

\end{tabular}
\end{center}
	\caption{Decidability results for VASS.}
\end{figure}

\begin{proposition}
{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
\end{proposition}

\begin{proof}
Consider a given $k \in \N$,
 $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
consider a given vector addition system $V$.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.

The set $\Bad$ admits a finite basis $B_\Bad$.
Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
and thus, are in $\Bad$, by upward-closedness.

Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
\end{proof}

\iffalse
\mathieu{
resilience never hold in the case of VAS. 
What happens when you add states ? \\
So you consider for instance $q_0(u_k)$ instead of just $u_k$. \\
Everything reachable from $q_0(u_k)$ is still of the form $q(u)$ with u greater than $v_{\Bad}$
(we'd have $v_{\Bad}$ sum of the coordinate of the elements from the basis of $\Bad$ when not considering states).
So now, let us fix some state $q$.
Either there is an element of the basis of $\Bad$ with state $q$.
Then $q(u)$ is necessarily in $\Bad$ (If this hold for all $q \in Q$, then again $k$-resilience does not hold).
Either there is none.
If there is none, then $q(\omega)$ is in the complement of $\Bad$, i.e. $\Safe$.
Then, of course there are probably other elements in $\Safe$ but from
the perspective of $\Bad$ and the extremely high elements $q_i(u_k)$  these ‘smaller’ elements are all but unreachable. 
And the question becomes rather `can we reach $q(\omega)$ from every element of $\Bad$ save from maybe the few smaller ones ?’
Since $q(\omega)$ here is upward-closed this is decidable i.e. it is asking coverability questions. 
It’s not 100\% the question «asking, for every element in the basis of $\Bad$, whether or not $q(0)$ is coverable from it», since you could have, for smaller elements in $\Bad$, other elements of $\Safe$ you can reach, but it seem close.
This would lead to a decision scheme.
Need to write this more formally/properly.}
\fi

\begin{theorem}\label{brp VASS}
{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
.
\end{theorem}

\begin{proof}
\mathieu{Donner l'idée et se référencer à l'appendix~\ref{appendix brp VASS} pour la preuve complête.}
\end{proof}