### 2023-06-13 abstract.tex
up

We investigate and expand upon the notion of resilience as defined in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state. We introduce a more general definition of resilience and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
and the $k$-(state-)resilience problem (kRP). The (general) resilience problem 
% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$. 
We remark that it is an instance of the home space problem defined in the framework of VASS.
State resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.

In a well-structured framework we show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility. Moreover, state-resilience,
bounded-state-resilience and
 $k$-state-resilience
are undecidable for strongly upward-compatible WSTS with effective pred-basis.

On the positive side, we find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$, and moreover the resilience problem is decidable for ideal-effective WSTS with 
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$
and
the additional hypothesis that
for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.

We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

Finally, we study the resilience for VASS, where most problems are shown decidable.