### 2023-06-21 abstract.tex
section abstract

\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-state property) defined by the two subsets Bad and Safe and by the existence of a bound on the length of runs starting from Bad and reaching Safe. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility and with effective pred-basis. Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility. Moreover, some resilience properties are decidable for two other classes of WSTS : (1) for WSTS with effective 
$\uparrow$ $\post^*$ basis and (2) for ideal-effective WSTS with downward and upward compatibilities. Finally, we study the resilience for VASS where most of the problems are shown decidable.