### 2023-06-21 abstract.tex
section abstract

\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-space property) defined by Bad is an upward/downward closed set or it is a subset of the reachability set from an initial state, Safe is upward-closed ($\Safe=\uparrow \Safe$) or Safe is downward-closed ($\Safe=\downarrow \Safe$) (Bad is the complementary of Safe) and by the existence (or not) of a bound $k$ on the length of runs starting from Bad and reaching Safe. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis). Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$. Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective 
$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$. Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for extensions and variations of VASS; most of the resilience properties are shown decidable.